### 2020-10-07 perihelion_advance_of_mercury.Rmd
more work, mostly clarifying motivation for the methods used

---
title: "Perihelion advance of Mercury"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("mercury.R")
source("usefulfuncs.R")
source("usefulgeodesicfuncs.R")
```

## Summary

The perihelion advance of Mercury is a well-known test of general relativity.  Einstein calculated that Mercury's perihelion would advance by 43 seconds of arc per century due to general relativistic effects; here I reproduce that analysis using numerical reasoning.

## Introduction

The orbit of Mercury has long been known to advance at a rate of about
575'' [seconds of arc] per century.  Various classical phenomena, such
as the effects of Venus, and the oblateness of the sun, account for
all but 43'' and this remaining amount puzzled astronomers for many
years.  According to Einstein and others, the general theory of
relativity predicts an additional precession of almost exactly 43
seconds per century, thus closing the gap between expectation and
observation.  In this document I will calculate the precessional rate
using numerical methods.

## Closed and non-closed orbits.

Consider the following R idiom, in which `trajectory()` is used to
directly simulate a geodesic in Schwarzschild space:

```{r,label=geodesicsimulatorGR}
# Simulate a trajectory starting at r=30, mass of sun=1/2 so Schwarzschild radius=1:
jj <- trajectory(r_start=30, m=1/2, h=3) # h=3 gives angular momentum
plot(jj$x,jj$y, asp=1, type='l', axes=FALSE, xlab='', ylab='',main='Geodesics in the Schwarzschild geometry')
polargrid(r=10*(1:3))
event_horizon()  # show the event horizon, to scale
```

(NB `trajectory()` takes geometrized units).  In the above, we see
a typical trajectory in Schwarzschild space with a precession of
around 40 degrees per orbit.  As a consistency check we can perform
the same calculation but with Newtonian approximations:

```{r,label=geodesicsimulatorclassical}
jj <- trajectory(30, 1/2, 3, GR=FALSE) # GR=FALSE means use Newtonian approximation
plot(jj$x,jj$y, asp=1, type='l', axes=FALSE, xlab='', ylab='',main='Newtonian approximation')
polargrid(r=10*(1:3))
event_horizon()
```

and above we see that the orbit is accurately closed, as expected
(three orbits are simulated, the lines overlap).  However, a more
exacting test would be to plot the same lines but focussing in on the
aphelion point $(30,0)$:

```{r,label=closeupofaphelion}
dx <- 0.01
dy <- 0.5
plot(jj$x,jj$y,xlim=c(29.99,30.001),ylim=c(-dy,dy),type='b',xlab='x',ylab='y')
```

In the above we see an accurately closed orbit (which simulates more
than three full revolutions round the sun).

## Mercury

First we specify some orbital parameters for the orbit of Mercury, mostly taken from Wikipedia:

```{r,label=mercuryorbitalparameters}
G <-  6.674e-11     # big G in SI
M  <- 1.9885e30     # mass of Sun (kg)
M_merc <- 3.3011e23 # mass of Mercury (kg)
sol <- 299792458    # speed of light (m/s)
merc_perihelion <- 46001200e3 
merc_aphelion   <- 69816900e3
mu <- G*(M + M_merc)
```

Now we calculate some features of the orbit in geometrized units:

```{r,label=geometrizedunits} 
d <- G*M/sol^2  # mass of sun in meters
r_s <- 2*d      # Schwarzschild radius of Sun (=2M)
```

And some further features using classical geometry:

```{r,label=furtherfeaturesclassical}
a <- (merc_perihelion + merc_aphelion)/2
cee <- a-merc_perihelion
b <- sqrt(a^2-cee^2)
ecc <- sqrt(1-(b/a)^2)  # should match eccentricity given by Wikipedia
ell <- (a^2-cee^2)/a  # semi-latus rectum
```

Now calculate maximum and minimum orbital speeds:

```{r,label=maxandminorbitalperiods}
vmin <- sqrt(mu/ell)*(1-ecc)
vmax <- sqrt(mu/ell)*(1+ecc)
```

And we need to specify the orbital angular mometum in different units:

```{r,label=orbitalangularmomentumSI}   
specific_angular_momentum_SI <- merc_perihelion * vmax
h <- specific_angular_momentum_SI/sol #geometrized units
```

With all these we can simulate the orbit:
```{r,label=orbitsimulatorincludingGR,cache=TRUE}
 out <- trajectory(merc_perihelion, m=d, h=h, tau=seq(from=0,to=35e15,len=5e5))
 plot(out$x,out$y,type='l',asp=1,axes=FALSE,xlab='',ylab='',main='Orbit of Mercury, Sun not to scale')
 points(0,0,pch=21,bg='yellow',col='orange') # Schwarzschild radius invisible at this scale
 max(out$phi/(2*pi))
```

In the above diagram we see over 15.4 orbits of Mercury, exhibiting no
visible deviation from an exactly closed orbital path.  Nevertheless,
the orbit is not closed as Mercury has a tiny amount of orbital
precession.  Note that the Schwarzschild radius would be microsopic at
this scale (the size of the central disk is dominated by the width of
the plotting line used for the event horizon).  To calculate this
precession we need to simulate the motion for a longer time:

```{r,label=orbitsimulatorGRlongtime,cache=TRUE}
out <- trajectory(merc_perihelion,m=d,h=h,tau=seq(from=0,to=35e16,len=1e7)) # takes ~5 mins
tail(out)
```

From the above we see that the trajectory has been simulated up to
about $965$ radians or about $\frac{965}{2\pi}\simeq 153$ revolutions
round the sun.  Plotting the entire trajectory is computationally
expensive, but we can plot the simulated trajectory close to aphelion
as for the Newtonian case:

```{r, label-plotmercuryorbitGR}
maxx <- 46001200027
n <- 650000
s <- seq_len(n)
plot(out$x[s],out$y[s],xlim=c(maxx-500,maxx+100),ylim=c(-5e6,5e6),type='o',xlab="x",ylab="y",pch=c(16,rep(1,n-1)))
```

In the above plot, note the differing axis scales.  The start point is
shown as a solid circle, subseqent points shown with open circles.
Lines connect consecutive points in the simulation.  It is difficult
to see evidence for precession in this plot because the points
correspond to regular snapshots of Mercury's position in successive
orbits, and the time evolution is not shown.  To proceed we need to
find the aphelion (or an approximation to it) somehow, and show that
the aphelion systematically increases with time.  Given the numerics,
we find points on the simulated trajectory which are a local maximum
distance from the sun.  This is straightfoward numerically, we
identify iterations $i$ (which have radius `r[i]`) satisfying both
`r[i-1]<r[i]` and `r[i]>r[i+1]`:

```{r,label=orbitalangleatperihelion}
jj1 <- c(NA,out$rdot)
jj2 <- c(out$rdot,NA)
change <- (jj1>0) & (jj2<0)

tee    <- out$time[change]

phi <- out$phi[change]
phi <- (phi) %% (2*pi)

plot(tee,phi-pi,main='orbital angle of Mercury at perihelion',xlab='time (seconds)',ylab='angle of perihelion')
```

The above diagram shows a clear trend, somewhat obscured by a vertical
spread due to the discrete nature of the simulation.  To estimate the
underlying trend, which is the general relativistic precession of
Mercury, we can use linear regression:

```{r,label=calculateprecession}
prec <- lm(phi~tee)$coefficients[2]
prec
```

And to convert to astronomical units we use

```{r,label=converttoastronomialunits}
conversion <-
  c(
      seconds_per_minute = 60,
      minutes_per_hour   = 60,
      hours_per_day      = 24,
      days_per_year      = 365.25,
      years_per_century  = 100,
      degrees_per_radian = 180/pi,
      minutes_per_degree = 60,
      seconds_per_minute = 60
  )
```

giving

```{r,label=resultofprecessionconverted}
prec*prod(conversion)
```

closely agreeing with Einstein's original prediction of 43''.  A more
careful analysis would account for the nonzero mass of Mercury.

By way of a consistency check, we can plot a similar diagram but with
the classical approximation:

```{r,precessionwithclassicaldynamics,cache=TRUE}
out <- trajectory(merc_perihelion, m=d, h=h, tau=seq(from=0,to=35e16,len=1e7),GR=FALSE)
# trajectory() takes ~ 5 mins to run
jj1 <- c(NA,out$rdot)
jj2 <- c(out$rdot,NA)
change <- (jj1>0) & (jj2<0)
tee    <- out$time[change]
phi <- out$phi[change]
phi <- (phi) %% (2*pi)
plot(tee,phi-pi,main='orbital angle at of Mercury at perihelion, classical case',xlab='time (seconds)',ylab='angle of perihelion')
```

In the above plot, note how there is no trend, consistent with the
Newtonian case of zero perihelion advance.

##  Theoretical analysis

In current notation, a first-order analysis gives

\[
\frac{24\pi^3a^2}{t^2c^2\left(1-e^2\right)}
\]

for the precession.  In R idiom:

```{r,label=theoreticalprecessioncalculator}
precession <- 24*pi^3*a^2/(per^2*sol^2*(1-ecc^2))  # radians/orbit
precession
```

To convert this figure (which is given in radians per orbit) to the
more usual units of arcseconds per century, we need to multiply by a
sequence of factors:

```{r,label=converttohumancentricunits}
seconds_per_orbit <- prod(c(
    seconds_per_minute = 60,
    minutes_per_hour   = 60,
    hours_per_day      = 24,
    days_per_orbit     = 87.969
))

seconds_per_century <- prod(c(
    seconds_per_minute = 60,
    minutes_per_hour   = 60,
    hours_per_day      = 24,
    days_per_year      = 365.25,
    years_per_century  = 100
))

orbits_per_century <- seconds_per_century/seconds_per_orbit

arcseconds_per_radian <- prod(
    c(
        degrees_per_radian = 180/pi,
        minutes_per_degree = 60,
        seconds_per_minute = 60
   	     ))

arcseconds_per_orbit <- precession*arcseconds_per_radian
arcseconds_per_century <- arcseconds_per_orbit * orbits_per_century
arcseconds_per_century
```

Compare with the value of about 43.37 obtained from numerical
integration, showing close agreement.