### 2021-01-25 physics_educator_rebuttal.txt
issue #22

Reviewers' comments:

Reviewer 1: This paper studies geodesics on the equatorial plane of
the spatial sections (constant time slices) of the Schwarzschild
spacetime. The geodesic equation in this case reduces to a second
order non-linear ODE, given by Eq. (6) in the paper, and it is solved
numerically. A few geodesics are plotted in Fig. 1, 2, 3, and 5.

The author's idea is to use the path of light taut strings to
visualize "straight lines in the vicinity of a black hole". The paper
claims that light taut strings will follow paths that minimize proper
length, so the paths are geodesics in spatial sections of the
Schwarzschild spacetime in the second paragraph in the introduction
section and also in the discussion section. However, this major claim
was made without any justification. Justification for this major claim
needs to be given.

The paper has imprecise usage of GR terminology. A few statements are
unclear. Examples include:

1.  "straight line" in the vicinity of black hole... (Introduction/paragraph 1)
I believe it refers to straight line in spatial sections of black hole.

2. large (but finite) amounts of string are required to... (second line after Eq. 3)
why large amounts are needed since w is small?

3. with radial strings being arbitrarily longer than flat space would
allow... (below Eq. 4) what is the meaning of this statement?

4. corresponding equation for null geodesics... (second line below Eq. 6)
I believe the author refers to the orbit equation of null geodesics confined to the equatorial plane.

5. phi' should be r' (page 4 line 5)

6. Local observer (page 4 line 7)
I believe the author means static observer in Schwarzschild coordinates.

7. Radius of curvature (page 4 line 7) Since the equatorial plane is
now curved, what radius of curvature is the author talking about?
Clarification is needed.

8. A radius of \sqrt{27M} (page 4 line 9) What radius is this?

9. Although photon paths provide a definition of
straightness... (discussion section line 2) In GR, worldlines of
photon are null geodesics. The paper does not distinguish geodesics of
spatial sections and geodesics of spacetime.