### 2021-01-27 physics_educator_rebuttal.txt
fixes #32

Rebuttal follows.  Reviewers' comments preceded by a chevron (">") and
my response included inline.

Overall response: the comments are prefectly reasonable and I have
addressed them individually.  The manuscript is much stengthened as a
result and I commend it to you.

> Reviewers' comments:
>  
> Reviewer 1: This paper studies geodesics on the equatorial plane of
> the spatial sections (constant time slices) of the Schwarzschild
> spacetime. The geodesic equation in this case reduces to a second
> order non-linear ODE, given by Eq. (6) in the paper, and it is solved
> numerically. A few geodesics are plotted in Fig. 1, 2, 3, and 5.
>  
> The author's idea is to use the path of light taut strings to
> visualize "straight lines in the vicinity of a black hole". The paper
> claims that light taut strings will follow paths that minimize proper
> length, so the paths are geodesics in spatial sections of the
> Schwarzschild spacetime in the second paragraph in the introduction
> section and also in the discussion section. However, this major claim
> was made without any justification. Justification for this major claim
> needs to be given.
>

On the one hand, this is a perfectly reasonable request: the claim is
indeed a major one, and one that had no explicit justification in the
manuscript.  But on the other, it requires no justification as it is
true by defnition.  Didactically, one might conceptualise a string as
being comprised of many tiny links, each of which contributes some
small length to the string, and each of which is sufficiently small
for Euclidean geometry to describe its length.  Each link is subject
to a well-defined local tension with respect to its neighbours.

We may then *count* the links and use this as a proxy for proper
length, demonstrating the consistency of the concept of proper length;
and of course small links oblige one to use infinitesimal calculus.

Imagine two strings, S1 and S2, which both smoothly join point p to
point q (perhaps in a Schwarzschild space), and their paths are close
in the sense of the Calculus of Variations.  Further, imagine that S1
has a proper length slightly exceeding that of S2.  If we then impose
a tension on the strings, we see that the S1 is unstable with respect
to S2, because S2 has fewer links.  Thus equilibrium paths of light
inextensible strings are at least local minima of length.

However, much of the above reasoning is necessarily tautological; it
is perfectly reasonable to *define* an "inextensible light string
under tension" as one that adopts a minimal [or at least extremal]
proper-length path between its end points, and i have included a short
discussion in the manuscript.

> The paper has imprecise usage of GR terminology. A few statements are
> unclear. Examples include:
>  
> 1.  "straight line" in the vicinity of black
> hole... (Introduction/paragraph 1) I believe it refers to straight
> line in spatial sections of black hole.
>

The original sentence was:

:: The sentence was: Here I show how considering light inextensible 
:: string can be used to interpret the concept of "straight line" in the
:: vicinity of a black hole.

It might be rephrased: "...interpret the concept of "straight line" in
spatial sections of a black hole", but I believe this form of words
would be potentially misleading.  My thinking was that the concept of
"straight line" ---in the abstract---is to be re-interpreted in the
manuscript as the path taken by a light inextensible string.

It now reads "Here I interpret a straight line as the path adopted by
a stationary, light inexensible string under tension passing between
two points"

There is a little bit of repetition in the revised manuscript, but I
think it is worth it for the extra clarity.

> 2. large (but finite) amounts of string are required to... (second
> line after Eq. 3) why large amounts are needed since w is small?

That is an extremely good observation.  The word "large" is
inappropriate and misleading here, I think.  Asymptotic expansions can
be confusing and what I meant to say was that the ratio of proper
string length to Euclidean string length (that is, length as estimated
by the Euclidean formula (1+r)-1 where 'r' is circumferential radius)
approaches infinity.  But the word "large" is not right here, and I
have rephrased accordingly.

I have also realised that this equation and the preceding one for
small w can be made much more comprehensible by giving an example, and
I have included two examples using solar-mass black holes.

>  
> 3. with radial strings being arbitrarily longer than flat space would
> allow... (below Eq. 4) what is the meaning of this statement?
>

Euclidean geometry predicts that l=2Mw; the intent of that sentence
was to point out that we Euclidean predictions are arbirtarily
incorrect as we move away from the black hole.  I have rephrased.

    NOT YET FINALISED THIS

> 4. corresponding equation for null geodesics... (second line below
> Eq. 6) I believe the author refers to the orbit equation of null
> geodesics confined to the equatorial plane.

corrected.

>  
> 5. phi' should be r' (page 4 line 5)
>

Manuscript now corrected.

> 6. Local observer (page 4 line 7) I believe the author means static
> observer in Schwarzschild coordinates.

I realise that I have been misusing the word "local" to mean "static"
and have now corrected the manuscript.

>  
> 7. Radius of curvature (page 4 line 7) Since the equatorial plane is
> now curved, what radius of curvature is the author talking about?
> Clarification is needed.

On re-reading, the original manuscript was a little cryptic.  I have
added a detailed derivation of this result that should clarify the
issue.  In essence, a light beam and a tangential string will diverge
quadratically and this leads to the radius of curvature quoted.

>  
> 8. A radius of \sqrt{27M} (page 4 line 9) What radius is this?

Derivation added as a result of point 7 above should clarify this
issue.

>  
> 9. Although photon paths provide a definition of
> straightness... (discussion section line 2) In GR, worldlines of
> photon are null geodesics. The paper does not distinguish geodesics of
> spatial sections and geodesics of spacetime.
>

I would argue that emphasising the concept of "spatial sections" might
be somewhat counterproductive, even though I admittedly present a
diagram of Flamm's paraboloid [which is obligatory here, I would say].
A spatial section S is a 3D subset of a 4D spacetime in which we
require the squared interval between any two distinct events of S to
be spacelike.  It is easy to give an example of a spatial section in
flat Minkowski space: with standard coordinates (t,x,y,z) we consider
events (t0,x,y,z) for some fixed t0.  Such constructions are somewhat
more difficult in Schwarzschild space but we might use Schwarzschild
coordinates (t,r,theta,phi) and then specify S2=(t0,r,theta,phi).  The
Schwarzschild metric then shows that S2 is a spatial section, provided
that r>2M, because the Schwarzschild metric (equation 1 in the
manuscript) is diagonal, and the nonvanishing components are strictly
positive.

When considering static strings it is easier and more convenient to
use standard Schwarzschild coordinates (t,r,theta,phi) and then simply
ignore the 't' coordinate as it plays no part in calculating proper
length of a string (being careful to specify that we are outside the
event horizon).  If the string is static, we may imagine two static
observers tied to the string; in calculating the amount of proper
distance between the two observers (string length) we simply ignore
the 't' coordinate and integrate the spatial component of the
Schwarzschild metric.  This process is distinct from the concept of
spatial section in which we condition on t=t0.

The world lines of freely-falling objects follow non-null geodesics
and photons follow null geodesics.  Light inextensible string follows
neither of these paths.  I think the revised manuscript makes this
clear