### 2021-02-24 tpe_edreview_rebuttal.txt
start addressing issue #22

Review of "Light inextensible strings (thread) under tension in the
Schwarzschild geometry" submitted to The Physics Educator

I want to begin by congratulating the author on a very clever
exercise.  If the route of a non-stretchable string under tension
would take the path of least distance, asking about the behavior of a
tight string in a Schwarzschild geometry, and comparing that path to
that of a light ray, offers an interesting question.  One might
initially suppose the two paths would be the same, or approximately
so.  The author rightly observes that General Relativity is a "topic
of long-standing fascination for physics undergraduates." True, but I
am unsure whether this particular exercise will offer more
understanding or more confusion for the intended undergraduates,
because the distinction between a stretched string and a null metrics
may be blurred (see point 11 below).

I would point out that at least one calculation presented in the paper
can found in existing introductory General Relativity textbooks.
Specifically, the meaning of the r-coordinate, and the difference
between Delta r and the distance between two concentric shells of
different r-coordinates (eq. 2) is thoroughly discussed in Taylor &
Wheeler's book Exploring Black Holes: Introduction to General
Relativity (Addison-Wesley-Longman, 2000), Ch. 2.  While the
calculation of Eq. (2) should be included in this paper for
completeness and done before going on to other calculations, Taylor &
Wheeler's black holes book is a standard pedagogical reference for
such calculations, and should be included among the references.
Taylor & Wheeler's book is very likely the intended undergraduate
student's first encounter with General Relativity.

The paper raises a few questions in this reviewer's mind, and in
elsewhere it offers opportunities for suggestions.

1.  Question.  First paragraph in "Introduction": The third sentence
"Here I interpret straight line as the path adopted by a stationary,
light inextensible string under tension passing between two points."
Is this to be understood as identical to the definition of "geodesic"?
If that is the assumption, can it be justified, because when one
thinks of geodesic in General Relativity, one has "free-fall" in mind,
and the stretched string, being "stationary," would not be in
free-fall in the global coordinate system.

2.  Suggestion about the line that begins on p. 1 and continues to
p. 2, about "lengthscales $L\leq T/(\mu g)$: Saying it this way may
present a sticking point for the undergraduate audience towards whom
this paper in intended.  The logic behind this inequality would show
greater clarity if one says something like "the weight of the string
does not exceed the tension, so that $L\mu g<T$".  From this the
stated inequality about lengthscales follows easily.

3.  Suggestion.  Stating "$c^4/4GM$ for the strength of gravity at the
event horizon" may leave an undergraduate audience baffled.  By
starting from where they are, with $g=GM/r^2$, then setting r equal to
the Schwarzschild radius, the stated result follows immediately,
without mystery.

4.  Question. Lines 3-4, page 2.  Can the author explain what is meant
by "the dominant energy condition"?

5.  Suggestion.  2nd and 3rd line below eq. (2): the statement "A
circular string at constant r will, trivially, have proper length
$2\pi r$."  Stating the relation this way muddies the logic and the
distinction between flat and Schwarzschild space.  The r-coordinate of
a spherical shell centered on a point mass M is defined operationally
by measuring the circumference of the shell and dividing the
circumference's value by $2\pi$.  That must be understood before the
calculations of Eq. (2) can make sense.  As the sentence is presently
structured it may obscure the distinction between the r coordinate and
radial lengths because it invites the undergraduate reader to
interpret the "triviality" as a statement about plane Euclidean
geometry instead of the Schwarzschild geometry's definition of the
r-coordinate.

6.  Question.  There may be a typographical error in Eq. (2).
Shouldn't the coefficient of the logarithm be 2M instead of 1/2?  As
written the equation is not dimensionally consistent, since r and M
are measured in the same units.  Compare to Taylor & Wheeler's
Exploring Black Holes, p. 2-28, Eq. (27).

7.  Suggestion: It appears that the $L$ and the $l$ that appear
between Eq. (1) and Eq. (3) denote the same quantity.  Why use
different symbols?

8.  Suggestion: In the line above Eq. (3), if $w >> 0$ then one does
not need absolute value bars in $|w|<< 1$.

9.  Question.  2nd line below Eq. (4), why are radial strings
"arbitrarily" longer than flat space would imply?  It appears the
difference is a matter of calculation (or measurement), not arbitrary.
Please explain this choice of words.

10. Suggestion.  Line below Eq. (5), I think the word "dashes" should
be "primes".  This is a dash: --- and this is a prime: ', at least in
usage with which I am familiar.

11.  The paragraph below Eq. (6) is puzzling.  Here the path of the
tight string is being compared to an "equatorial null geodesic," which
I take to mean the path of a light beam moving along the equator.  If
we are considering a light beam, then shouldn't a term with $t$ in it
appear somewhere?  The puzzlement continues given the statement that
"light, being sensitive to the time component of the metric" whereas
the equations discussed in this paragraph say nothing (that I can see)
about the time component of the metric.  The proper time interval
would be zero, but the coordinate time interval $dt$ would, I think,
be nonzero.  What am I missing?

12.  In the passage below Eq. (6), I am having trouble making the
connection between $r''(\phi)=r-3M + 2(r')^2/r$ and the claim that the
solution is $r=r_0\sec(\phi-\phi_0)$.  What trig identity am I
missing?

13.  Can the author explain more fully what is meant by a "tangential
null geodesic" (several lines after Eq. 6)?  That terminology makes it
sound like the r-coordinate does not change, whence $r'$ would be
zero, but that appears to not be the case.  In this section, if $r'$
is small I see where the approximation $r''=r-2M$ comes from, but I am
having trouble seeing where $r''-r-3M$ arises for the alleged
tangential null geodesics.  What am I not seeing?

I will have to take the author's word for the numerical results.

This is an interesting paper.  I congratulate the author for thinking
of such an intriguing application of the Schwarzschild metric.
However, in this reviewer's opinion, before the paper can be well
fitted to its announced audience ("an undergraduate homework problem")
some points need to be clarified.