### 2021-02-24 tpe_edreview_rebuttal.txt
fixes #40

Herewith my rebuttal to the review of "Light inextensible strings..."
The referee's comments are preceded by a chevron (">") and my response
is inline.  I have tinkered with linebreaks here and there in the
interests of clarity, and have replaced some mathematical notation in
the original review with latex equivalents.

Short story: I agree with all the comments made and have modified the
manuscript accordingly; changes appear in the manuscript using colour
coding for ease of reference.  The document is stronger, more
cohesive, more accurate, and hopefully more instructive; and I commend
it to you.

> Review of "Light inextensible strings (thread) under tension in the
> Schwarzschild geometry" submitted to The Physics Educator
> 
> I want to begin by congratulating the author on a very clever
> exercise.  If the route of a non-stretchable string under tension
> would take the path of least distance, asking about the behavior of a
> tight string in a Schwarzschild geometry, and comparing that path to
> that of a light ray, offers an interesting question.  One might
> initially suppose the two paths would be the same, or approximately
> so.  The author rightly observes that General Relativity is a "topic
> of long-standing fascination for physics undergraduates." True, but I
> am unsure whether this particular exercise will offer more
> understanding or more confusion for the intended undergraduates,
> because the distinction between a stretched string and a null metrics
> may be blurred (see point 11 below).
> 
> I would point out that at least one calculation presented in the paper
> can found in existing introductory General Relativity textbooks.
> Specifically, the meaning of the r-coordinate, and the difference
> between Delta r and the distance between two concentric shells of
> different r-coordinates (eq. 2) is thoroughly discussed in Taylor &
> Wheeler's book Exploring Black Holes: Introduction to General
> Relativity (Addison-Wesley-Longman, 2000), Ch. 2.  While the
> calculation of Eq. (2) should be included in this paper for
> completeness and done before going on to other calculations, Taylor &
> Wheeler's black holes book is a standard pedagogical reference for
> such calculations, and should be included among the references.
> Taylor & Wheeler's book is very likely the intended undergraduate
> student's first encounter with General Relativity.
> 
> The paper raises a few questions in this reviewer's mind, and in
> elsewhere it offers opportunities for suggestions.
> 
> 1.  Question.  First paragraph in "Introduction": The third sentence
> "Here I interpret straight line as the path adopted by a stationary,
> light inextensible string under tension passing between two points."
> Is this to be understood as identical to the definition of "geodesic"?
> If that is the assumption, can it be justified, because when one
> thinks of geodesic in General Relativity, one has "free-fall" in mind,
> and the stretched string, being "stationary," would not be in
> free-fall in the global coordinate system.
> 
> 2.  Suggestion about the line that begins on p. 1 and continues to
> p. 2, about "lengthscales $L\leq T/(\mu g)$: Saying it this way may
> present a sticking point for the undergraduate audience towards whom
> this paper in intended.  The logic behind this inequality would show
> greater clarity if one says something like "the weight of the string
> does not exceed the tension, so that $L\mu g<T$".  From this the
> stated inequality about lengthscales follows easily.
> 
> 3.  Suggestion.  Stating "$c^4/4GM$ for the strength of gravity at the
> event horizon" may leave an undergraduate audience baffled.  By
> starting from where they are, with $g=GM/r^2$, then setting r equal to
> the Schwarzschild radius, the stated result follows immediately,
> without mystery.
> 
> 4.  Question. Lines 3-4, page 2.  Can the author explain what is meant
> by "the dominant energy condition"?
> 
> 5.  Suggestion.  2nd and 3rd line below eq. (2): the statement "A
> circular string at constant r will, trivially, have proper length
> $2\pi r$."  Stating the relation this way muddies the logic and the
> distinction between flat and Schwarzschild space.  The r-coordinate of
> a spherical shell centered on a point mass M is defined operationally
> by measuring the circumference of the shell and dividing the
> circumference's value by $2\pi$.  That must be understood before the
> calculations of Eq. (2) can make sense.  As the sentence is presently
> structured it may obscure the distinction between the r coordinate and
> radial lengths because it invites the undergraduate reader to
> interpret the "triviality" as a statement about plane Euclidean
> geometry instead of the Schwarzschild geometry's definition of the
> r-coordinate.
> 
> 6.  Question.  There may be a typographical error in Eq. (2).
> Shouldn't the coefficient of the logarithm be 2M instead of 1/2?  As
> written the equation is not dimensionally consistent, since r and M
> are measured in the same units.  Compare to Taylor & Wheeler's
> Exploring Black Holes, p. 2-28, Eq. (27).
> 
> 7.  Suggestion: It appears that the $L$ and the $l$ that appear
> between Eq. (1) and Eq. (3) denote the same quantity.  Why use
> different symbols?
>

corrected, now using $\ell$ consistently for proper length.

> 8.  Suggestion: In the line above Eq. (3), if $w >> 0$ then one does
> not need absolute value bars in $|w|<< 1$.
> 
> 9.  Question.  2nd line below Eq. (4), why are radial strings
> "arbitrarily" longer than flat space would imply?  It appears the
> difference is a matter of calculation (or measurement), not arbitrary.
> Please explain this choice of words.
> 
> 10. Suggestion.  Line below Eq. (5), I think the word "dashes" should
> be "primes".  This is a dash: --- and this is a prime: ', at least in
> usage with which I am familiar.
> 
> 11.  The paragraph below Eq. (6) is puzzling.  Here the path of the
> tight string is being compared to an "equatorial null geodesic," which
> I take to mean the path of a light beam moving along the equator.  If
> we are considering a light beam, then shouldn't a term with $t$ in it
> appear somewhere?  The puzzlement continues given the statement that
> "light, being sensitive to the time component of the metric" whereas
> the equations discussed in this paragraph say nothing (that I can see)
> about the time component of the metric.  The proper time interval
> would be zero, but the coordinate time interval $dt$ would, I think,
> be nonzero.  What am I missing?
> 
> 12.  In the passage below Eq. (6), I am having trouble making the
> connection between $r''(\phi)=r-3M + 2(r')^2/r$ and the claim that the
> solution is $r=r_0\sec(\phi-\phi_0)$.  What trig identity am I
> missing?
> 
> 13.  Can the author explain more fully what is meant by a "tangential
> null geodesic" (several lines after Eq. 6)?  That terminology makes it
> sound like the r-coordinate does not change, whence $r'$ would be
> zero, but that appears to not be the case.  In this section, if $r'$
> is small I see where the approximation $r''=r-2M$ comes from, but I am
> having trouble seeing where $r''-r-3M$ arises for the alleged
> tangential null geodesics.  What am I not seeing?
> 
> I will have to take the author's word for the numerical results.
> 
> This is an interesting paper.  I congratulate the author for thinking
> of such an intriguing application of the Schwarzschild metric.
> However, in this reviewer's opinion, before the paper can be well
> fitted to its announced audience ("an undergraduate homework problem")
> some points need to be clarified.