### 2021-02-28 tpe_edreview_rebuttal.txt
further minor rewording

Herewith my rebuttal to the review of "Light inextensible strings..."
The referee's comments are preceded by a chevron (">") and my response
is inline.  I have tinkered with linebreaks here and there in the
interests of clarity, and have replaced some mathematical notation in
the original review with latex equivalents.

Short story: I agree with all the comments made and have modified the
manuscript accordingly.  The document is stronger, more cohesive, more
accurate, and hopefully more instructive.  I submit that the extra
clarity and pedagogical value more than compensates for the slight
extra length.

I have changed the discussion slightly (emphasising the
staticness of the metric considered) as a result of reflecting on the
totality of the comments too.

> Review of "Light inextensible strings (thread) under tension in the
> Schwarzschild geometry" submitted to The Physics Educator
> 
> I want to begin by congratulating the author on a very clever
> exercise.  If the route of a non-stretchable string under tension
> would take the path of least distance, asking about the behavior of a
> tight string in a Schwarzschild geometry, and comparing that path to
> that of a light ray, offers an interesting question.  One might
> initially suppose the two paths would be the same, or approximately
> so.  The author rightly observes that General Relativity is a "topic
> of long-standing fascination for physics undergraduates." True, but I
> am unsure whether this particular exercise will offer more
> understanding or more confusion for the intended undergraduates,
> because the distinction between a stretched string and a null metrics
> may be blurred (see point 11 below).
> 
> I would point out that at least one calculation presented in the paper
> can found in existing introductory General Relativity textbooks.
> Specifically, the meaning of the r-coordinate, and the difference
> between Delta r and the distance between two concentric shells of
> different r-coordinates (eq. 2) is thoroughly discussed in Taylor &
> Wheeler's book Exploring Black Holes: Introduction to General
> Relativity (Addison-Wesley-Longman, 2000), Ch. 2.  While the
> calculation of Eq. (2) should be included in this paper for
> completeness and done before going on to other calculations, Taylor &
> Wheeler's black holes book is a standard pedagogical reference for
> such calculations, and should be included among the references.
> Taylor & Wheeler's book is very likely the intended undergraduate
> student's first encounter with General Relativity.

I was unaware of the derivation in Taylor and Wheeler that the
reviewer points me to.  I did not want to suggest that equation 2 was
new, but was unable to find a derivation of it in the literature.  I
have cited Taylor and Wheeler (but see my response to Q6 below).

> 
> The paper raises a few questions in this reviewer's mind, and in
> elsewhere it offers opportunities for suggestions.
> 
> 1.  Question.  First paragraph in "Introduction": The third sentence
> "Here I interpret straight line as the path adopted by a stationary,
> light inextensible string under tension passing between two points."
> Is this to be understood as identical to the definition of "geodesic"?
> If that is the assumption, can it be justified, because when one
> thinks of geodesic in General Relativity, one has "free-fall" in mind,
> and the stretched string, being "stationary," would not be in
> free-fall in the global coordinate system.

I think the sentence in question was technically correct but
potentially misleading.  What I wanted to do was to make readers think
about taut thread in the context of GR.  I didn't want to consider
geodesics at this point in the introduction, or free-falling elevator
frames.  I want to focus on taut string.

[one of the conceptual difficulties in preparing the manuscript, as
pointed out by a previous reviewer, is that the string is not in a
freely-falling elevator frame; as the reviewer points out, such string
does not follow a geodesic path in any meaningful sense].

I have rephrased the sentence which now reads: "Here I consider the
path adopted by a stationary, light inexensible string under tension
passing between two fixed points, and consider the extent to which the
expression "straight line" would be applicable to such an object".

> 
> 2.  Suggestion about the line that begins on p. 1 and continues to
> p. 2, about "lengthscales $L\leq T/(\mu g)$: Saying it this way may
> present a sticking point for the undergraduate audience towards whom
> this paper in intended.  The logic behind this inequality would show
> greater clarity if one says something like "the weight of the string
> does not exceed the tension, so that $L\mu g<T$".  From this the
> stated inequality about lengthscales follows easily.
>

This is a much better way of phrasing the argument and I have
rephrased as suggested.

> 3.  Suggestion.  Stating "$c^4/4GM$ for the strength of gravity at the
> event horizon" may leave an undergraduate audience baffled.  By
> starting from where they are, with $g=GM/r^2$, then setting r equal to
> the Schwarzschild radius, the stated result follows immediately,
> without mystery.
>

I have totally rephrased that part of the manuscript using this
suggestion and it is much clearer (I have also changed uppercase L for
lowercase l as per suggestion number 7.

> 4.  Question. Lines 3-4, page 2.  Can the author explain what is meant
> by "the dominant energy condition"?

On reflection the dominant energy condition is not really an
appropriate or informative or useful or interesting concept for the
intended readership.  The important thing is that tension greater $\mu
c^2$ is unphysical and this context, the unphysicality is manifested
by transverse waves travelling faster than light.

In answer to the question, we may take the stress-energy tensor for a
stretched string to be diagonal with elements $\mu,-T,0,0$.  Then the
dominant energy condition is that, for any future-pointing causal
vector, $-T^\mu_\nu V^\nu$ must also be a future-pointing causal
vector.  We can see directly that if $T>\mu$ the DEC is violated.

But this is a massive digression to the main thrust of the manuscript
and any adequate pedagogical explanation would be an intrusive
distraction.  It is like using a slegehammer to crack a nut.  I have
removed any reference to the DEC on these grounds, and further
emphasised the unphysicality of the wave speed being $>c$.

> 
> 5.  Suggestion.  2nd and 3rd line below eq. (2): the statement "A
> circular string at constant r will, trivially, have proper length
> $2\pi r$."  Stating the relation this way muddies the logic and the
> distinction between flat and Schwarzschild space.  The r-coordinate of
> a spherical shell centered on a point mass M is defined operationally
> by measuring the circumference of the shell and dividing the
> circumference's value by $2\pi$.  That must be understood before the
> calculations of Eq. (2) can make sense.  As the sentence is presently
> structured it may obscure the distinction between the r coordinate and
> radial lengths because it invites the undergraduate reader to
> interpret the "triviality" as a statement about plane Euclidean
> geometry instead of the Schwarzschild geometry's definition of the
> r-coordinate.
>

The referee is absolutely correct here and raises an important point.
When I said that the circumference was "trivially" $2\pi r$, what I
meant was that integrating equation (1) from $phi=0$ to $2\pi$ at
fixed $t,r,\theta$ was trivial.  The resulting proper string length is
a direct consequence of the *definition* of "r" [which is sometimes
called the circumferential radial coordinate].

This comment has made me realise that the $2\pi r$ proper string
length is actually the simplest proper length calculation possible in
the Schwarzschild metric, and I have rephrased accordingly, and I hope
it is more instructive.  I remember struggling with this myself as an
undergraduate (I wanted to know what the "real" radius was ...).
Einstein himself said "it is not so easy to free oneself from the idea
that coordinates must have an immediate metrical meaning", so I have
good company!

> 6.  Question.  There may be a typographical error in Eq. (2).
> Shouldn't the coefficient of the logarithm be 2M instead of 1/2?  As
> written the equation is not dimensionally consistent, since r and M
> are measured in the same units.  Compare to Taylor & Wheeler's
> Exploring Black Holes, p. 2-28, Eq. (27).
>

Factor of M now added (actually this caused me some considerable
confusion and uncertainty.  The equation presented in Taylor and
Wheeler has a factor of 2M but it should be M---see the attached
mathematica screenshot, plus the identity arctanh(x) =
log((1+x)/(1-x))/2.  In the same section Taylor and Wheeler discuss
one-centimeter long rulers laid end to end many times, a good metaphor
for proper length.

This was just a typo in that particular line and has no effect on any
other part of the manuscript.

> 7.  Suggestion: It appears that the $L$ and the $l$ that appear
> between Eq. (1) and Eq. (3) denote the same quantity.  Why use
> different symbols?
>

corrected, now using $\ell$ consistently for proper length.

> 8.  Suggestion: In the line above Eq. (3), if $w >> 0$ then one does
> not need absolute value bars in $|w|<< 1$.
>

Fixed, now reads "consider the proper length from $r=2M$ to
$r=2M(1+w)$ where $w$ is small, $0<w\ll 1$".

> 9.  Question.  2nd line below Eq. (4), why are radial strings
> "arbitrarily" longer than flat space would imply?  It appears the
> difference is a matter of calculation (or measurement), not arbitrary.
> Please explain this choice of words.
>

In mathematical usage (but not in general usage, at least according to
the OED) there is a distinction between "arbitrary" and "arbitrarily".
One usually uses phrases like "arbitrarily large" to convey the idea
that some object may become as large as one desires.

In the manuscript the wording was "radial strings are arbitrarily
longer than flat space would imply", which I think gets the intended
meaning across.

In this case we have the startling result that the discrepancy
approaches infinity ("becomes arbitrarily large") as w approaches
infinity.  I have not seen this result explictly before, yet am
reluctant to claim that it is truly new: after all it is a simple
consequence of a known result (equation 3).  But it is certainly
startling, to me at least.

In any case the original wording was somewhat terse and I have
rephrased.  The manuscript now explicitly states that the discrepancy
in proper length between the Schwarzschild and Euclidean results
approaches infinity as w approaches infinity.

> 10. Suggestion.  Line below Eq. (5), I think the word "dashes" should
> be "primes".  This is a dash: --- and this is a prime: ', at least in
> usage with which I am familiar.

Fixed, notation now consistent

> 
> 11.  The paragraph below Eq. (6) is puzzling.  Here the path of the
> tight string is being compared to an "equatorial null geodesic," which
> I take to mean the path of a light beam moving along the equator.  If
> we are considering a light beam, then shouldn't a term with $t$ in it
> appear somewhere?  The puzzlement continues given the statement that
> "light, being sensitive to the time component of the metric" whereas
> the equations discussed in this paragraph say nothing (that I can see)
> about the time component of the metric.  The proper time interval
> would be zero, but the coordinate time interval $dt$ would, I think,
> be nonzero.  What am I missing?
>

The equation given as $u''=3u^2/2-u$ gives the path of a beam of light
through Schwarzschild space in which $r$ is given as a function of
angle, $r=r(\phi)$.  The idea is that a stationary observer has a
laser pointer pointing in a particular direction, switches it on and
leaves it on for a long time.  We then record all the points that the
beam touches.  The path of the light is then simply $r=r(\phi)$, no
time coordinate involved, as Figures 4 and 6 show.  I have changed the
citation to Rindler, who gives a detailed derivation.

> 12.  In the passage below Eq. (6), I am having trouble making the
> connection between $r''(\phi)=r-3M + 2(r')^2/r$ and the claim that the
> solution is $r=r_0\sec(\phi-\phi_0)$.  What trig identity am I
> missing?
>

There are two separate issues here.  Firstly, null geodesic paths in
the Schwarzschild metric obey $u''=3u^2/2-u$ where $u=2M/r$.  This is
equivalent to $r''(\phi) = r-3M + 2(r')^2/r$.  The second issue is
straight lines in Euclidean flat space (in flat space the expression
"straight lne" is unambiguous).  WLOG $\phi_0=0$ and if
$r=r_0\sec\phi$ then $r'=-r_0\sin\phi\cdot\sec^2\phi$ and
$r''=r_0(\sec\phi + 2\sin^2\phi\sec^3\phi$, from which $r''=2(r')^2/r$
follows from cancellation.

One can "go the other way"---that is, given ODE $r''=r+2(r')^2/r$, one
can deriving $r=r_0\sec\phi$---by substituting $u=1/r$ and obtaining
$u''+u=0$, whence $u=A\cos(\phi-\phi_0)$.

I have rephrased slightly and numbered one of the equations so I can
refer to it.

> 13.  Can the author explain more fully what is meant by a "tangential
> null geodesic" (several lines after Eq. 6)?  That terminology makes it
> sound like the r-coordinate does not change, whence $r'$ would be
> zero, but that appears to not be the case.  In this section, if $r'$
> is small I see where the approximation $r''=r-2M$ comes from, but I am
> having trouble seeing where $r''-r-3M$ arises for the alleged
> tangential null geodesics.  What am I not seeing?

Everything can be resolved and clarified by omitting the word
"tangential" which was misleading, and replacing "small" with
"negligible".  The sentence now reads "If $r'$ is negligible, then a
string follows $r''=r-2M$, while null geodesics follow $r''=r-3M$".

The $r''=r-3M$ is just the (new) equation 7 -viz
$r''\left(\phi\right)=r-3M+2\left(r'\right)^2/r$ but with $r'=0$.

> 
> I will have to take the author's word for the numerical results.
>

Not any more! All the software used to create the diagrams is now
publically available under version control at

https://github.com/RobinHankin/string

and I have added a citation to this repo.  The code is heavily
documented.

> This is an interesting paper.  I congratulate the author for thinking
> of such an intriguing application of the Schwarzschild metric.
> However, in this reviewer's opinion, before the paper can be well
> fitted to its announced audience ("an undergraduate homework problem")
> some points need to be clarified.