### 2024-10-31 weighted_string.Rmd
use Rmd format here

---
title: "Heavy inextensible strings in the Schwarzschild geometry"
author: "Robin Hankin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("deSolve")
```

Now define the  second derivative from Euler's equation:

```{r definerdashdash}
rdashdash <- function(r,rdash,lambda,M=1/2){
    x <- sqrt(1-2*M/r)
    return(
    (M*(4+lambda*(4-3*x))*r^2 + (-1+lambda*(-1+x))*r^3 + M*(3 + lambda*(3-2*x))*rdash^2 +
     2*r*(M^2*(-2+lambda*(-2+x)) + (-1+lambda*(-1+x))*rdash^2))/((-1+lambda*(-1+x))*r*(-2*M+r))
    )
}
```

```{r useode}
string <- function(r0,lambda=0,angles=seq(from = 0, to = 0.5, by = 0.01)){

    parameters <- c(lambda = lambda, M = 1/2)
    state <- c(r = r0, rdash = 0)
    times <- angles

    Lorenz <- function(t, state, parameters){
        with(as.list(c(state, parameters)), {
            dr <- rdash
            drdash <- rdashdash(r, rdash, lambda, M)
            list(c(dr, drdash))
        })
    }

    out <- ode(y = state, times = times, func = Lorenz, parms = parameters)

    phi <- out[,1]
    r <- out[,2]

    return(cbind(r*cos(phi),r*sin(phi)))
}
```

```{r plotstrings}

size <- 5
plot(c(-size,size),c(-size,size),type="n",asp=1)
th <- seq(from=0,to=2*pi,len=100)
points(cos(th),sin(th),type='l')  # event horizon
points(string(r0=1.4,lambda= 0,angles=seq(0,2,0.01)),type='l')
for(l in seq(from=-0.1,to=-1.4,by=-0.1)){
    points(string(r0=1.4,lambda=l,angles=seq(0,pi/2,0.01)),type='l')
}
```